Exoplanet population inference and the abundance \\ of Earth analogs from noisy, incomplete catalogs

### 2014-08-21 document/ms.tex
adjusting wording of intro discussion of inv-det-eff

There are two qualitatively different methods that are commonly used to infer
the occurrence rate density from a catalog and a detection efficiency
specification.
The first is an intuitive method that we will refer to as
``inverse-detection-efficiency'' and the second is based on the likelihood
function of the catalog given a parametric rate density.
The inverse-detection-efficiency method involves making a histogram of the
objects in the catalog where each point is weighted by its inverse detection
probability.
This method is very popular in the literature (\citealt{howard, dong,
dressing, swift, petigura}) even though it is not motivated probabilistically.
The alternative likelihood method models the catalog as a Poisson realization
of the \emph{observable} rate density of exoplanets taking the survey
detection efficiencies and transit probabilities into account.
This technique has been used to constrain parametric models---a broken power
law, for example---for the occurrence rate density (\citealt{tabachnik,
youdin, dong}).
As we show in \sect{model}, if we model the occurrence rate density as a
piecewise constant function, the inverse-detection-efficiency method can be
derived as a special case of the likelihood method with a specific set of
assumptions.

@@

\acknowledgments
We would like to thank Erik Petigura (Berkeley) for freely sharing his data
and code.
It is a pleasure to thank
Ruth Angus (Oxford),
Tom Barclay (NASA Ames),
Jo Bovy (IAS),
Eric Ford (PSU), and
Ben Montet (Caltech/Harvard)
for helpful contributions to the ideas and code presented here.
We would also like to acknowlege the anonymous referee for suggestions that
substantially improved the paper.
This project was partially supported by the NSF (grant AST-0908357), NASA
(grant NNX08AJ48G), and the Moore--Sloan Data Science Environment at NYU.
This research builds on ideas generated at a three-week workshop supported by
NSF Grant DMS-1127914 to the Statistical and Applied Mathematical Sciences
Institute.
This research made use of the NASA \project{Astrophysics Data System}.

@@

\section{Inverse-detection-efficiency}

>>>>

There are two qualitatively different methods that are commonly used to infer
the occurrence rate density from a catalog and a detection efficiency
specification.
The first is an intuitive method that we will refer to as
``inverse-detection-efficiency'' and the second is based on the likelihood
function of the catalog given a parametric rate density.
The inverse-detection-efficiency method involves making a histogram of the
objects in the catalog where each point is weighted by its inverse detection
probability.
This method is very popular in the literature (\citealt{howard, dong,
dressing, swift, petigura}) even though it is not motivated probabilistically.
The alternative likelihood method models the catalog as a Poisson realization
of the \emph{observable} rate density of exoplanets taking the survey
detection efficiencies and transit probabilities into account.
This technique has been used to constrain parametric models---a broken power
law, for example---for the occurrence rate density (\citealt{tabachnik,
youdin, dong}).
In this \paper, we start from the likelihood method but model the rate density
non-parametrically as a piecewise-constant step function.
Using this formulation of the problem, we derive a generalization that takes
observational uncertainties into account.
In \app{inv-det-eff}, we show that the inverse-detection-efficiency method can
be derived as a special case of the likelihood method in the limit of a
smoothly varying completeness function.

@@

\acknowledgments
We would like to thank Erik Petigura (Berkeley) for freely sharing his data
and code.
It is a pleasure to thank
Ruth Angus (Oxford),
Tom Barclay (NASA Ames),
Jo Bovy (IAS),
Eric Ford (PSU),
David Kipping (CfA),
Ben Montet (Caltech/Harvard), and
Scott Tremaine (IAS)
for helpful contributions to the ideas and code presented here.
We would also like to acknowlege the anonymous referee for suggestions that
substantially improved the paper.
This project was partially supported by the NSF (grant AST-0908357), NASA
(grant NNX08AJ48G), and the Moore--Sloan Data Science Environment at NYU.
This research builds on ideas generated at a three-week workshop supported by
NSF Grant DMS-1127914 to the Statistical and Applied Mathematical Sciences
Institute.
This research made use of the NASA \project{Astrophysics Data System}.

@@

\section{Inverse-detection-efficiency}
\sectlabel{inv-det-eff}
